---
title: "individual1"
author: "Neha Maddali"
date: "12/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
```

```{r}
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(lubridate)
library(reshape2)
```

```{r warning=FALSE}
#import the data
data <- readxl::read_xlsx('ClimateChange.xlsx', sheet='Data')
country <- readxl::read_xlsx('ClimateChange.xlsx', sheet='Country')
series <- readxl::read_xlsx('ClimateChange.xlsx', sheet='Series')

#change the ".." cells in the years to n/a or 0
data <- na_if(data, '..')
```


**4.) How does population growth affect GDP?**

**Variable utilized:**
<br />
SP.POP.GROW = Population growth (annual %)
<br />
NY.GDP.MKTP.CD = GDP($)
```{r warning=FALSE}
pop <- data %>% filter(`Series code` == 'SP.POP.TOTL') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

pop<-melt(pop, id.vars=c("Country name", "Series code"))

gdp <- data %>% filter(`Series code` == 'NY.GDP.MKTP.CD') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

gdp<-melt(gdp, id.vars=c("Country name", "Series code"))

test <- pop %>% inner_join(gdp %>% select("Country name",`Series code`,`variable`,`value`), 
                           by=c("Country name","variable"))
test <- na.omit(test)                           # Apply na.omit function


data1 <- test %>% group_by(`Country name`) %>%
  summarize(`Country name`, "year" = `variable`, "population" = `value.x`, "GDP"=`value.y`)

datSig <- filter(data1, `Country name` %in% c("China", "Germany", "India", "Iran, Islamic Rep.", 
                                            "Japan", "Korea, Dem. Rep.", "Russian Federation", 
                                            "United States"))
ggplot(datSig, aes(x=population, y=GDP, colour = `Country name`, group = `Country name`)) + geom_point()+ geom_line() +labs(x="Population Growth (annual %)", y="GDP ($)", title="GDP Vs. Population Growth (annual %) from 1990-2011")
```


**5.) Which country has the most environmental impact from climate change in the time period 1990-2011?**

**Variable utilized:**
<br />
EN.CLC.MDAT.ZS = Droughts, floods, extreme temps (% pop. avg. 1990-2009)
<br />
ER.H2O.FWTL.ZS = Annual freshwater withdrawals (% of internal resources)
<br />
AG.LND.EL5M.ZS = Land area below 5m (% of land area)

```{r warning=FALSE}
disaster <- data %>% filter(`Series code` == 'EN.CLC.MDAT.ZS') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)
disaster<-melt(disaster, id.vars=c("Country name", "Series code"))

disaster <- filter(disaster, `Country name` %in% c("China", "Germany", "India", 
                                                  "Iran, Islamic Rep.", "Japan", 
                                                  "Korea, Dem. Rep.", "Russian Federation", 
                                                  "United States"))
h2oWithdraw <- data %>% filter(`Series code` == 'ER.H2O.FWTL.ZS') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)
h2oWithdraw<-melt(h2oWithdraw, id.vars=c("Country name", "Series code"))

h2oWithdraw <- filter(h2oWithdraw, `Country name` %in% c("China", "Germany", "India", 
                                                   "Iran, Islamic Rep.", "Japan", 
                                                   "Korea, Dem. Rep.", "Russian Federation", 
                                                   "United States"))
landBelow <- data %>% filter(`Series code` == 'AG.LND.EL5M.ZS') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)
landBelow<-melt(landBelow, id.vars=c("Country name", "Series code"))

landBelow <- filter(landBelow, `Country name` %in% c("China", "Germany", "India", 
                                                         "Iran, Islamic Rep.", "Japan", 
                                                         "Korea, Dem. Rep.", "Russian Federation", 
                                                         "United States"))

impact <- disaster %>% inner_join(h2oWithdraw %>% select("Country name",`Series code`,`variable`,`value`), 
                           by=c("Country name","variable"))
impact <- impact %>% inner_join(landBelow %>% select("Country name",`Series code`,`variable`,`value`), 
                                  by=c("Country name","variable"))
impact <- impact %>% group_by(`Country name`) %>%
  summarize(`Country name`, "year" = `variable`, "Natural Disasters" = `value.x`, 
            "H20 Withdrawal"=`value.y`, "Land Below"=`value`) %>% 
  mutate(`Natural Disasters` = as.numeric(`Natural Disasters`)) %>% 
  mutate(`H20 Withdrawal` = as.numeric(`H20 Withdrawal`)) %>% 
  mutate(`Land Below` = as.numeric(`Land Below`))

#test <- impact %>% summarize(`Country name`, `Natural Disasters`, `H20 Withdrawal`, `Land Below`)

impactChina <- filter(impact, `Country name` %in% c("China"))
a <- mean(impactChina$`Natural Disasters`,na.rm=TRUE)
a1 <- mean(impactChina$`H20 Withdrawal`,na.rm=TRUE)
a2 <- mean(impactChina$`Land Below`,na.rm=TRUE)

impactGermany <- filter(impact, `Country name` %in% c("Germany"))
b <- mean(impactGermany$`Natural Disasters`,na.rm=TRUE)
b1 <- mean(impactGermany$`H20 Withdrawal`,na.rm=TRUE)
b2 <- mean(impactGermany$`Land Below`,na.rm=TRUE)

impactIndia <- filter(impact, `Country name` %in% c("India"))
c <- mean(impactIndia$`Natural Disasters`,na.rm=TRUE)
c1 <- mean(impactIndia$`H20 Withdrawal`,na.rm=TRUE)
c2 <- mean(impactIndia$`Land Below`,na.rm=TRUE)

impactUS <- filter(impact, `Country name` %in% c("United States"))
d <- mean(impactUS$`Natural Disasters`,na.rm=TRUE)
d1 <- mean(impactUS$`H20 Withdrawal`,na.rm=TRUE)
d2 <- mean(impactUS$`Land Below`,na.rm=TRUE)

impactJapan <- filter(impact, `Country name` %in% c("Japan"))
e <- mean(impactJapan$`Natural Disasters`,na.rm=TRUE)
e1 <- mean(impactJapan$`H20 Withdrawal`,na.rm=TRUE)
e2 <- mean(impactJapan$`Land Below`,na.rm=TRUE)

impactRussia <- filter(impact, `Country name` %in% c("Russian Federation"))
f <- mean(impactRussia$`Natural Disasters`,na.rm=TRUE)
f1 <- mean(impactRussia$`H20 Withdrawal`,na.rm=TRUE)
f2 <- mean(impactRussia$`Land Below`,na.rm=TRUE)

impactKorea <- filter(impact, `Country name` %in% c("Korea, Dem. Rep."))
g <- mean(impactKorea$`Natural Disasters`,na.rm=TRUE)
g1 <- mean(impactKorea$`H20 Withdrawal`,na.rm=TRUE)
g2 <- mean(impactKorea$`Land Below`,na.rm=TRUE)

impactIran <- filter(impact, `Country name` %in% c("Iran, Islamic Rep."))
h <- mean(impactIran$`Natural Disasters`,na.rm=TRUE)
h1 <- mean(impactIran$`H20 Withdrawal`,na.rm=TRUE)
h2 <- mean(impactIran$`Land Below`,na.rm=TRUE)

country <- c('China','Germany','India','United States','Japan','Russian Federation','Korea, Dem. Rep.','Iran, Islamic Rep.')
avgNaturalDisasters <- c(a,b,c,d,e,f,g,h)
avgH20Withdrawal <- c(a1,b1,c1,d1,e1,f1,g1,h1)
avgLandBelow <- c(a2,b2,c2,d2,e2,f2,g2,h2)

envirImpact <- data.frame(country, avgNaturalDisasters, avgH20Withdrawal, avgLandBelow)
envirImpact <- envirImpact %>% rowwise() %>% 
  mutate(total = sum(c(avgNaturalDisasters, avgH20Withdrawal, avgLandBelow))) %>% 
  summarize(country, avgNaturalDisasters, avgH20Withdrawal, avgLandBelow, total)

lbls <- c('China','Germany','India','United States','Japan','Russian Federation','Korea, Dem. Rep.','Iran, Islamic Rep.')
pct <- round(envirImpact$total/sum(envirImpact$total)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie(envirImpact$total,labels = lbls, col=rainbow(length(lbls)),
    main="Country impact by Climate Change")
```



**6.) How does GDP affect amount of gas emissions?**

**Variable utilized:**
EN.ATM.CO2E.KT = CO2 emissions, total (KtCO2)
NY.GDP.MKTP.CD = GDP($)

```{r}
gas <- data %>% filter(`Series code` == 'EN.ATM.CO2E.KT') %>% 
  summarize(Country_Name =`Country name`, Series_Code =`Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)
gas<-melt(gas, id.vars=c("Country_Name", "Series_Code"))
gas <- gas %>%
  group_by(Country_Name) %>%
  summarize(Country_Name,Series_Code, year=`variable`,gas_emission=`value`)
gas$Series_Code <- NULL

gassignificant <- filter(gas, Country_Name %in% c("China", "Germany", "India", 
                                                  "Iran, Islamic Rep.", "Japan", 
                                                  "Korea, Dem. Rep.", "Russian Federation", 
                                                  "United States"))

gdp <- data %>% filter(`Series code` == 'NY.GDP.MKTP.CD') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)
gdp<-melt(gdp, id.vars=c("Country name", "Series code"))
gdp <- gdp %>%
  group_by(`Country name`) %>%
  summarize(Country_Name=`Country name`,`Series code`, year=`variable`,GDP=`value`)
gdp$`Series code` <- NULL
gdp$`Country name` <- NULL

gdpSig <- filter(gdp, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

effect <- gdpSig %>%
  inner_join(gassignificant %>%
               select(Country_Name, year, gas_emission),
             by=c("Country_Name", "year"))
effect <- na.omit(effect)                           # Apply na.omit function

#multiple graphs in one figure
ggplot(effect, aes(x=gas_emission, y=GDP)) + geom_point() + facet_wrap(~Country_Name, scales="free") + 
  labs(x="Gas Emissions (KtCO2)", y="GDP", title="GDP Vs. Gas Emissions from 1990-2011") + 
  theme(axis.text.x = element_text(angle = 90)) + geom_smooth(method="lm") 
```
