---
title: "EDA Pt2"
author: "Neha Maddali"
date: "2024-02-25"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(sf)
library(tigris)
library(mapview)
library(tidycensus)
library(leaflet)
library(tmap)
library(ggplot2)
library(plotly)
library(shiny)
```

```{r}
cfs <- read_rds("C:/Users/neham/Desktop/DS401/DS 401 Spring 2024/Source Data/CGC_CFS_2018-2023.rds")
```

### Manipulating dataframe
Grouped by jurisdiction and cfs_type, count the occurrences, for exploration.

```{r}
cfs_type_per_jurisdiction <- cfs %>%
  group_by(jurisdiction, cfs_type) %>%
  summarise(count = n())
#View(cfs_type_per_jurisdiction)
```

Create a new column that categorizes 'Weekday' or 'Weekend'
```{r}
cfs <- cfs %>%
  mutate(weekdays = ifelse(cfs_weekday %in% 2:6, "Weekday", "Weekend"))
```


### Trends
Excluding Other category, since a lot of the data is classified as "Other". Can look more closely at substance related categories 
```{r, echo=FALSE, warning=FALSE}
filtered_cfs <- cfs %>%
  filter(category != "Other")

call_source_category_counts <- filtered_cfs %>%
  group_by(jurisdiction, category) %>%
  summarise(count = n()) %>%
  arrange(jurisdiction, desc(count))

ggplot(call_source_category_counts, aes(x = reorder(jurisdiction, -count), y = count, fill = category)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Calls from Different Sources by Category (Excluding 'Other')",
       x = "Call Source",
       y = "Count",
       fill = "Call Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_viridis_d()  # Adjust the color scale as needed
```

```{r}
# Count occurrences of each call source
call_source_counts <- cfs %>%
  group_by(jurisdiction) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

ggplot(call_source_category_counts, aes(x = reorder(jurisdiction, -count), y = count / sum(count), fill = category)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of Calls from Different Sources by Category (Excluding 'Other')",
       x = "Call Source",
       y = "Percentage",
       fill = "Call Category") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_viridis_d()  # Adjust the color scale as needed
```


This plot looks at the monthly distribution of calls by weekday
```{r}
ggplot(cfs, aes(x = cfs_month, fill = factor(cfs_weekday))) +
  geom_bar(position = "stack") +
  labs(title = "Monthly Distribution of Calls by Weekday",
       x = "Month",
       y = "Count",
       fill = "Weekday") +
  theme_minimal()
```

Similar plot looking at the distribution of calls by category and weekday
```{r}
ggplot(cfs, aes(x = weekdays, fill = category)) +
  geom_bar(position = "stack") +
  labs(title = "Distribution of Calls by Category and Weekday",
       x = "Weekday",
       y = "Count",
       fill = "Category") +
  theme_minimal()
```

Filter the categories column for substance-related categories. Plot that shows the distribution of substance-related calls by weekday.
```{r}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories)

ggplot(substance_cfs, aes(x = factor(cfs_weekday), fill = category)) +
  geom_bar(position = "stack") +
  labs(title = "Distribution of Substance-Related Calls by Weekday",
       x = "Weekday",
       y = "Count",
       fill = "Category") +
  scale_x_discrete(labels = c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat")) +
  theme_minimal()
```


```{r}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories)

substance_cfs_counts <- substance_cfs %>%
  group_by(cfs_weekday, category) %>%
  summarise(count = n()) %>%
  arrange(cfs_weekday, desc(count))

substance_cfs_percentages <- substance_cfs_counts %>%
  group_by(cfs_weekday) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(substance_cfs_percentages, aes(x = factor(cfs_weekday), y = percentage, fill = category)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of Substance-Related Calls by Weekday",
       x = "Weekday",
       y = "Percentage",
       fill = "Category") +
  scale_x_discrete(labels = c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat")) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()
```


Distribute across the 3 PD's
```{r}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories)

# Create a plot with facets for each combination of weekday and call source
ggplot(substance_cfs, aes(x = factor(cfs_weekday), fill = category)) +
  geom_bar(position = "stack") +
  facet_grid(. ~ jurisdiction) +
  labs(title = "Distribution of Substance-Related Calls by Weekday and Call Source",
       x = "Weekday",
       y = "Count",
       fill = "Category") +
  scale_x_discrete(labels = c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat")) +
  theme_minimal()
```

```{r}
substance_cfs_counts <- substance_cfs %>%
  group_by(jurisdiction, cfs_weekday, category) %>%
  summarise(count = n()) %>%
  arrange(cfs_weekday, jurisdiction, desc(count))

substance_cfs_proportions <- substance_cfs_counts %>%
  group_by(jurisdiction, cfs_weekday) %>%
  mutate(proportion = count / sum(count))

ggplot(substance_cfs_proportions, aes(x = factor(cfs_weekday), y = proportion, fill = category)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_grid(. ~ jurisdiction) +
  labs(title = "Distribution of Substance-Related Calls by Weekday and Call Source",
       x = "Weekday",
       y = "Proportion",
       fill = "Category") +
  scale_x_discrete(labels = c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat")) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Time series
```{r}
cfs$cfs_date <- as.Date(cfs$cfs_date)

cfs$year <- lubridate::year(cfs$cfs_date)
cfs$month <- lubridate::month(cfs$cfs_date)

# Create a time series plot for calls over months with lines for each year
ggplot(cfs, aes(x = month, y = ..count.., group = year, color = as.factor(year))) +
  geom_line(stat = "count", binwidth = 30) +  # Binwidth set to approximately one month
  labs(title = "Calls Over Months by Year",
       x = "Month",
       y = "Number of Calls",
       color = "Year") +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Show month abbreviations on x-axis
  theme_minimal()
```

```{r}
ggplot(cfs, aes(x = cfs_hour, group = year, color = as.factor(year))) +
  geom_freqpoly(binwidth = 1, size = 1) +  # Binwidth set to 1 hour
  labs(title = "Calls Over Hours of the Day by Year",
       x = "Hour of the Day",
       y = "Number of Calls",
       color = "Year") +
  theme_minimal() +
  scale_x_continuous(limits = c(1, 23), breaks = seq(1, 23, by = 1))

```

```{r}
# Convert cfs_date to a Date object if it's not already
cfs$cfs_date <- as.Date(cfs$cfs_date)

# Extract year and month from cfs_date
cfs$year <- lubridate::year(cfs$cfs_date)
cfs$month <- lubridate::month(cfs$cfs_date)

# Filter out "Other" category
filtered_cfs <- cfs %>%
  filter(!(category %in% c("Other", "Nuisance", "Health", "Blight", "Domestic")))

# Create an interactive time series plot with animation and facets
p <- ggplot(filtered_cfs, aes(x = cfs_hour, color = category, group = interaction(category, year))) +
  geom_freqpoly(binwidth = 1, size = 1) +
  labs(title = "Temporal Trends in Calls Over Hours of the Day by Category",
       x = "Hour of the Day",
       y = "Number of Calls",
       color = "Category") +
  theme_minimal() +
  theme(legend.position = "top") +
  facet_wrap(~year, scales = 'free_y', ncol = 2)

# Convert ggplot to plotly with animation
p <- ggplotly(p, animation_frame = ~year)
p
```

### Spatial
Across all years, substance related categories
```{r}
gb <- readRDS("C:/Users/neham/Desktop/DS401/DS 401 Spring 2024/Source Data/CGC_block_groups.rds")
cfspre <- readRDS("C:/Users/neham/Desktop/DS401/DS 401 Spring 2024/Source Data/CGC_CFS_2018-2023.rds")
grid <-readRDS("C:/Users/neham/Desktop/DS401/DS 401 Spring 2024/Source Data/CGC_grid.rds")
cfs <- st_drop_geometry(cfspre)
coordinates <- st_as_sf(cfs, coords = c("longitude", "latitude"), crs = 4326)


# Filter substance-related calls
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories)

# Create a spatial object for substance-related calls
substance_coordinates <- st_as_sf(substance_cfs, coords = c("longitude", "latitude"), crs = 4326)

# Count occurrences in grid squares
substance_counts <- substance_coordinates %>%
  st_drop_geometry() %>%
  group_by(grid_id) %>%
  summarise(count = n())

# Merge with grid data
grid_substance <- left_join(grid, substance_counts, by = "grid_id")

# Create map
mapview(grid_substance, zcol = "count", legend = TRUE)
```


For year 2018
```{r, echo=FALSE, results='hide',error=FALSE, warning=FALSE, message=FALSE}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories, cfs_year == 2018)
substance_coordinates <- st_as_sf(substance_cfs, coords = c("longitude", "latitude"), crs = 4326)

substance_counts <- substance_coordinates %>%
  st_drop_geometry() %>%
  group_by(grid_id) %>%
  summarise(count = n())
grid_substance <- left_join(grid, substance_counts, by = "grid_id")
```
```{r}
mapview(grid_substance, zcol = "count", legend = TRUE)
```

For year 2019
```{r, echo=FALSE, results='hide',error=FALSE, warning=FALSE, message=FALSE}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories, cfs_year == 2019)
substance_coordinates <- st_as_sf(substance_cfs, coords = c("longitude", "latitude"), crs = 4326)
substance_counts <- substance_coordinates %>%
  st_drop_geometry() %>%
  group_by(grid_id) %>%
  summarise(count = n())
grid_substance <- left_join(grid, substance_counts, by = "grid_id")
```
```{r}
mapview(grid_substance, zcol = "count", legend = TRUE)
```

For year 2020
```{r, echo=FALSE, results='hide',error=FALSE, warning=FALSE, message=FALSE}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories, cfs_year == 2020)
substance_coordinates <- st_as_sf(substance_cfs, coords = c("longitude", "latitude"), crs = 4326)
substance_counts <- substance_coordinates %>%
  st_drop_geometry() %>%
  group_by(grid_id) %>%
  summarise(count = n())
grid_substance <- left_join(grid, substance_counts, by = "grid_id")
```
```{r}
mapview(grid_substance, zcol = "count", legend = TRUE)
```

For year 2021
```{r, echo=FALSE, results='hide',error=FALSE, warning=FALSE, message=FALSE}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories, cfs_year == 2021)
substance_coordinates <- st_as_sf(substance_cfs, coords = c("longitude", "latitude"), crs = 4326)
substance_counts <- substance_coordinates %>%
  st_drop_geometry() %>%
  group_by(grid_id) %>%
  summarise(count = n())
grid_substance <- left_join(grid, substance_counts, by = "grid_id")
```
```{r}
mapview(grid_substance, zcol = "count", legend = TRUE)
```

For year 2022
```{r, echo=FALSE, results='hide',error=FALSE, warning=FALSE, message=FALSE}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories, cfs_year == 2022)
substance_coordinates <- st_as_sf(substance_cfs, coords = c("longitude", "latitude"), crs = 4326)
substance_counts <- substance_coordinates %>%
  st_drop_geometry() %>%
  group_by(grid_id) %>%
  summarise(count = n())
grid_substance <- left_join(grid, substance_counts, by = "grid_id")
```
```{r}
mapview(grid_substance, zcol = "count", legend = TRUE)
```

For year 2023
```{r, echo=FALSE, results='hide',error=FALSE, warning=FALSE, message=FALSE}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories, cfs_year == 2023)
substance_coordinates <- st_as_sf(substance_cfs, coords = c("longitude", "latitude"), crs = 4326)
substance_counts <- substance_coordinates %>%
  st_drop_geometry() %>%
  group_by(grid_id) %>%
  summarise(count = n())
grid_substance <- left_join(grid, substance_counts, by = "grid_id")
```
```{r}
mapview(grid_substance, zcol = "count", legend = TRUE)
```


### R shiny
```{r}
cfs <- readRDS("C:/Users/neham/Desktop/DS401/DS 401 Spring 2024/Source Data/CGC_CFS_2018-2023.rds")
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>% filter(category %in% substance_categories)

ui <- fluidPage(
  titlePanel("Temporal and Spatial Interaction for Substance-Related Calls"),
  sliderInput("hourSlider", "Select Hour:", min = 0, max = 23, value = 0, step = 1),
  leafletOutput("map")
)

server <- function(input, output) {
  # Filter data based on selected hour
  filtered_data <- reactive({
    substance_cfs %>%
      filter(cfs_hour == input$hourSlider)
  })

  category_colors <- c("#66c2a5", "#fc8d62", "#8da0cb")

  # Create map
  output$map <- renderLeaflet({
    leaflet() %>%
      addTiles() %>%
      addCircleMarkers(data = filtered_data(),
                       ~longitude, ~latitude,
                       popup = paste("Hour: ", input$hourSlider),
                       label = ~category,
                       color = ~factor(category, levels = substance_categories, labels = category_colors),
                       fillOpacity = 0.8)
  })
}

shinyApp(ui, server)
```

```{r}
# Load your data
cfs <- readRDS("C:/Users/neham/Desktop/DS401/DS 401 Spring 2024/Source Data/CGC_CFS_2018-2023.rds")
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>% filter(category %in% substance_categories)

# Define UI
ui <- fluidPage(
  titlePanel("Temporal and Spatial Interaction for Substance-Related Calls"),
  sliderInput("yearSlider", "Select Year:", min = 2018, max = 2023, value = 2018, step = 1),
  leafletOutput("map")
)

# Define server
server <- function(input, output) {
  # Filter data based on selected year
  filtered_data <- reactive({
    substance_cfs %>%
      filter(cfs_year == input$yearSlider)
  })

  category_colors <- c("#66c2a5", "#fc8d62", "#8da0cb")


  # Create map
  output$map <- renderLeaflet({
    leaflet() %>%
      addTiles() %>%
      addCircleMarkers(data = filtered_data(),
                       ~longitude, ~latitude,
                       popup = paste("Year: ", input$yearSlider),
                       label = ~category,
                       color = ~factor(category, levels = substance_categories, labels = category_colors),
                       fillOpacity = 0.8)
  })
}

# Run the application
shinyApp(ui, server)
```


### Logistic Regression
```{r}
cfs$substance_related <- ifelse(cfs$category %in% substance_categories, 1, 0)

logistic_model <- glm(substance_related ~ as.factor(cfs_weekday), data = cfs, family = "binomial")
summary(logistic_model)

odds_ratios <- exp(coef(logistic_model))
odds_ratios
```
