---
title: "Exploratory Data Analysis"
author: "Neha Maddali"
date: "2024-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(sf)
library(tigris)
library(mapview)
library(tidycensus)
library(leaflet)
library(tmap)
library(ggplot2)
```

## Exploring Census Data

```{r}
# Read tidy census data 
v17 <- load_variables(2017, "acs5", cache=TRUE)
```

The following looks at the distribution of sex (male/female) across the counties in Iowa.
```{r, results='hide',fig.keep='all', warning=FALSE}
# Get geospatial data -----------------------------------------------------
# Add sex-related variables to the list of variables
vars <- c(
  "B01001_001", # Total Population

  # Sex
  "B01001_002", # Male Population
  "B01001_026"  # Female Population
)

# Retrieve data with sex-related variables
counties <- get_acs(
  geography = "county", state = "IA",
  variables = vars, survey = "acs5",
  output = "wide", geometry = TRUE
) |>
  mutate(
    male_pct = B01001_002E / B01001_001E,
    female_pct = B01001_026E / B01001_001E
  ) |>
  st_transform(
    crs = 4326
  )

# Choropleth map for male population percentage using ggplot
ggplot(data = counties, aes(fill = male_pct)) +
  geom_sf() +
  scale_fill_viridis_c(option = "B", labels = scales::label_percent()) +
  labs(fill = "Percent\nMale Population") +
  theme_void()

# Choropleth map for female population percentage using ggplot
ggplot(data = counties, aes(fill = female_pct)) +
  geom_sf() +
  scale_fill_viridis_c(option = "B", labels = scales::label_percent()) +
  labs(fill = "Percent\nFemale Population") +
  theme_void()

```

The following looks at the distribution of age across the counties in Iowa.
```{r, results='hide',fig.keep='all'}
# Add age-related variables to the list of variables
vars <- c(
  "B01001_001", # Total Population

  # Age
  "B01001_003", "B01001_004", "B01001_005", "B01001_006", "B01001_007", "B01001_027", "B01001_028", "B01001_029", "B01001_030", "B01001_031"
)

# Retrieve data with age-related variables
counties <- get_acs(
  geography = "county", state = "IA",
  variables = vars, survey = "acs5",
  output = "wide", geometry = TRUE
) |>
  mutate(
    age_0_9_pct = (B01001_003E + B01001_004E + B01001_005E + B01001_006E + B01001_007E) / B01001_001E,
    age_25_34_pct = (B01001_027E + B01001_028E + B01001_029E + B01001_030E + B01001_031E) / B01001_001E
  ) |>
  st_transform(
    crs = 4326
  )

# Choropleth map for age 0-9 population percentage using ggplot
ggplot(data = counties, aes(fill = age_0_9_pct)) +
  geom_sf() +
  scale_fill_viridis_c(option = "B", labels = scales::label_percent()) +
  labs(fill = "Percent\nAge 0-9 Population") +
  theme_void()

# Choropleth map for age 25-34 population percentage using ggplot
ggplot(data = counties, aes(fill = age_25_34_pct)) +
  geom_sf() +
  scale_fill_viridis_c(option = "B", labels = scales::label_percent()) +
  labs(fill = "Percent\nAge 25-34 Population") +
  theme_void()
```



## Exploring CGC_CFS_2018-2023 Data

The following is from the geospatial_data.R example code file.
```{r, echo=FALSE, results='hide',error=FALSE, warning=FALSE, message=FALSE}
# Read CFS data ----------------------------------------------------
cfs <- read_rds("C:/Users/neham/Desktop/DS401/DS 401 Spring 2024/Source Data/CGC_CFS_2018-2023.rds")
class(cfs)
cfs_no_geo <- cfs |>
  sf::st_drop_geometry() # Remove geometry from data

class(cfs_no_geo)
# No longer "sf" type, which can make it easier to manipulate if you aren't creating maps or doing geospatial operations

# Get Place data ----------------------------------------------------------
ia_places <- places(state = "IA", year = 2020) |>
  st_transform(crs = 3417)

# Add place data to CFS ----------------------------------------------------------
cfs <- st_transform(cfs, crs = st_crs(ia_places))

# st_join performs a "spatial join", adding place data to the CFS data for calls that are located within a city
cfs_places <- st_join(cfs, ia_places)
```
```{r}
mapview(ia_places) + mapview(cfs_places |> sample_n(10000), zcol = "NAME")
```


### Manipulating dataframe
Grouped by jurisdiction and cfs_type, count the occurrences, for exploration.

```{r}
cfs_type_per_jurisdiction <- cfs %>%
  group_by(jurisdiction, cfs_type) %>%
  summarise(count = n())
#View(cfs_type_per_jurisdiction)
```

Create a new column that categorizes 'Weekday' or 'Weekend'
```{r}
cfs <- cfs %>%
  mutate(weekdays = ifelse(cfs_weekday %in% 2:6, "Weekday", "Weekend"))
```


### Summary stats for numerical variables
```{r}
summary(cfs[, c("cfs_year", "cfs_month", "cfs_hour", "cfs_minute")])

# Descriptive Statistics
# descriptive_stats <- cfs %>%
#   summarize(
#     Mean_cfs_year = mean(cfs_year),
#     Median_cfs_year = median(cfs_year),
#     Mode_cfs_year = as.numeric(names(table(cfs_year))[table(cfs_year) == max(table(cfs_year))]),
#     Max_cfs_year = max(cfs_year),
#     Min_cfs_year = min(cfs_year),
#     Q1_cfs_year = quantile(cfs_year, 0.25),
#     Q3_cfs_year = quantile(cfs_year, 0.75),
#     SD_cfs_year = sd(cfs_year)
#   )
# print(descriptive_stats)

```

### Categorical variable exploration
Identify the most frequent categories and jurisdiction called to
```{r}
# Count the occurrences of each jurisdiction
jurisdiction_counts <- cfs %>%
  group_by(jurisdiction) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
print(jurisdiction_counts)

# Count the occurrences of each category
category_counts <- cfs %>%
  group_by(category) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
print(category_counts)
```

Want to see what falls under "Other"
```{r}
other_calls <- cfs %>%
  filter(category == "Other")

unique_cfs_types_other <- unique(other_calls$cfs_type)
print(unique_cfs_types_other)

```

Looks like most of the data falls under Other
```{r}
# Count occurrences of each call type
call_type_counts <- cfs %>%
  group_by(category) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

ggplot(call_type_counts, aes(x = reorder(category, -count), y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Distribution of Call Types",
       x = "Call Type",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



### Trends
This plot looks at the distribution of calls from the 3 different sources
```{r}
# Count occurrences of each call source
call_source_counts <- cfs %>%
  group_by(jurisdiction) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

ggplot(call_source_counts, aes(x = reorder(jurisdiction, -count), y = count)) +
  geom_bar(stat = "identity", fill = "lightcoral") +
  labs(title = "Distribution of Calls from Different Sources",
       x = "Call Source",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The plot looks more in depth of what types of calls are being made across the 3 sources.
```{r}
# Count occurrences of each call source and category
call_source_category_counts <- cfs %>%
  group_by(jurisdiction, category) %>%
  summarise(count = n()) %>%
  arrange(jurisdiction, desc(count))

ggplot(call_source_category_counts, aes(x = reorder(jurisdiction, -count), y = count, fill = category)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Calls from Different Sources by Category",
       x = "Call Source",
       y = "Count",
       fill = "Call Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_viridis_d()
```

Excluding Other category, since a lot of the data is classified as "Other". Can look more closely at substance related categories 
```{r, echo=FALSE, warning=FALSE}
filtered_cfs <- cfs %>%
  filter(category != "Other")

call_source_category_counts <- filtered_cfs %>%
  group_by(jurisdiction, category) %>%
  summarise(count = n()) %>%
  arrange(jurisdiction, desc(count))

ggplot(call_source_category_counts, aes(x = reorder(jurisdiction, -count), y = count, fill = category)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Calls from Different Sources by Category (Excluding 'Other')",
       x = "Call Source",
       y = "Count",
       fill = "Call Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_viridis_d()  # Adjust the color scale as needed
```


This plot looks at the monthly distribution of calls by weekday
```{r}
ggplot(cfs, aes(x = cfs_month, fill = factor(cfs_weekday))) +
  geom_bar(position = "stack") +
  labs(title = "Monthly Distribution of Calls by Weekday",
       x = "Month",
       y = "Count",
       fill = "Weekday") +
  theme_minimal()
```

Similar plot looking at the distribution of calls by category and weekday
```{r}
ggplot(cfs, aes(x = weekdays, fill = category)) +
  geom_bar(position = "stack") +
  labs(title = "Distribution of Calls by Category and Weekday",
       x = "Weekday",
       y = "Count",
       fill = "Category") +
  theme_minimal()
```

Filter the categories column for substance-related categories. Plot that shows the distribution of substance-related calls by weekday.
```{r}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories)

ggplot(substance_cfs, aes(x = factor(cfs_weekday), fill = category)) +
  geom_bar(position = "stack") +
  labs(title = "Distribution of Substance-Related Calls by Weekday",
       x = "Weekday",
       y = "Count",
       fill = "Category") +
  scale_x_discrete(labels = c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat")) +
  theme_minimal()
```

Distribute across the 3 PD's
```{r}
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs %>%
  filter(category %in% substance_categories)

# Create a plot with facets for each combination of weekday and call source
ggplot(substance_cfs, aes(x = factor(cfs_weekday), fill = category)) +
  geom_bar(position = "stack") +
  facet_grid(. ~ jurisdiction) +
  labs(title = "Distribution of Substance-Related Calls by Weekday and Call Source",
       x = "Weekday",
       y = "Count",
       fill = "Category") +
  scale_x_discrete(labels = c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat")) +
  theme_minimal()
```


### Geospatial 
Focused on the substance related categories
```{r, results='hide',error=FALSE, warning=FALSE, message=FALSE}
ia_places <- places(state = "IA", year = 2020) |>
  st_transform(crs = 3417)

cfs <- st_transform(cfs, crs = st_crs(ia_places))
cfs_places <- st_join(cfs, ia_places)

# Filter substance-related calls
substance_categories <- c("Substance-On Premise", "Substance-Off Premise", "Substance-Driving")
substance_cfs <- cfs_places %>%
  filter(category %in% substance_categories)
```
```{r}
mapview(ia_places) + mapview(substance_cfs, zcol = "category")
```

