---
title: "hw5"
author: "Neha Maddali"
date: "11/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results='hide'}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(stringr)
library(lubridate) # Time
library(scales)
```

## Question 1
The complete dataset is too large. Download and import the alcohol sales data for Ames only:
```{r}
data <- read.csv('./2021_Iowa_Liquor_Sales.csv')
str(data)

new_data <- data
```

## Question 2
2a. Convert the Date column into a date object
```{r}
new_data <- new_data %>% mutate(Date = mdy(Date))
str(new_data)
```

2b. Extract the longitude and latitude from the Store Location column, creating two new columns
```{r}
new_data <- separate(new_data, col = "Store.Location",
                     into=c("Point","Latitude","Longitude"),sep="\\(")
new_data$Point <- NULL
new_data <- separate(new_data, col="Latitude", into=c("Latitude","Longitude"),sep="\\)")
new_data <- separate(new_data, col="Latitude", into=c("Latitude","Longitude"),sep=" ")
str(new_data)
```

## Question 3
First overview: Provide a visual breakdown of the liquor category (by Category Name). Include volume sold in the breakdown. Describe your figure. Which type of liquor is the most popular?
```{r}
new_data$`Category.Name` = factor(new_data$`Category.Name`)
temp_data <- new_data %>% arrange(desc(`Volume.Sold..Gallons.`))
head(temp_data)
labels = temp_data$`Category.Name`[1:20]

new_data %>% ggplot(aes(x = `Category.Name`), group_by = `Volume.Sold..Gallons.`) + 
  geom_bar() + 
  #scale_x_discrete(breaks = labels) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Category name", y = "Number sold")
```
<br />
ANALYSIS: American Vodkas is the most popular as it has the most number sold compared to the 
other categories.

## Question 4
Which day of the week has the most orders? Use a barchart to visualize the results. Explain the pattern and provide a reason why the pattern occurs.
```{r}
data1 <- new_data
data1$weekday <- weekdays(data1$Date)
str(data1)

data1 %>% ggplot(aes(x = weekday), group_by = `Volume.Sold..Gallons.`) + 
  geom_bar() +
  labs(x = "Days of the Week", y = "Number sold")
```
<br />

ANALYSIS: Thursdays have the most orders. This could be a preparation for the weekend type of 
parties that occur.

## Question 5
Create a single plot for the monthly volumes sold. Each liquor category should be shown using a separate time series. Label the top 5 categories using text labels.
```{r}
monthly <- new_data %>%
  group_by(Name=`Category.Name`, month=floor_date(new_data$Date, "month")) %>%
  summarize(totalVolumes = sum(`Volume.Sold..Gallons.`), Name, month)

monthlySummary <- monthly %>%
  group_by(Name) %>%
  summarize(month=max(month), totalVolumes = max(totalVolumes))
monthlySummary <- monthlySummary %>%
  arrange(desc(totalVolumes))

ggplot(monthly, aes(x=month, y=totalVolumes, by=Name))+geom_line() +
  labs(y="Total Volume Sold", title="Total Volume Sold vs. Months of Liquor") +
  geom_text(aes(x=month, y=totalVolumes, label=Name),data=monthly %>% filter(totalVolumes > 1210))
```
<br />

ANALYSIS: American Vodkas are the highest in volumes sold (gallons) per month. The top 5 categories are American Vodkas, Canadian Whiskies, Cocktails/RTD, Straight Bourbon Whiskies, and White Rum.

## Question 6
Which categories are spring graduation party favorites? Find at least one category.
```{r}
datTest <- new_data %>%
  select(Date, Volume.Sold..Gallons.,Category.Name,) %>%
  filter(between(Date, as.Date("2021-04-01"), as.Date("2021-06-30"))) %>%
  group_by(Name=`Category.Name`) %>%
  summarize(totalVolumes = sum(`Volume.Sold..Gallons.`)) %>%
  arrange(desc(totalVolumes))
str(datTest) 
```
<br />

ANALYSIS: For this problem, I looked at the volume sold in gallons for the months of April,
May, and June. American Vodkas is the highest category. The second and third are Canadian Whiskies, Cocktails/RTD. These can be the spring graduation party favorites.





